---
title: 'Homework #3'
author: "Hunter Ratliff"
date: "11/13/2019"
output:
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 2
  word_document: 
    highlight: tango
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)
library(pander)
library(tidyverse)
library(ggthemes)
library(scales)
options(tinytex.verbose = TRUE)

### LATEX HELP ###
# \definecolor{bgblue}{RGB}{217, 237, 247}
# \textcolor{red}{easily} \colorbox{bgblue}{this text}
# \colorbox{bgblue}{asd}
# \noindent
# {\color{bgblue} \rule{\linewidth}{1mm} }

# see http://joshua.smcvt.edu/latex2e/_005cfbox-_0026-_005cframebox.html
```

You can use SAS/R to solve these. Show the SAS/R code used to find the answers and be sure to write out hypothesis and the equations as appropriate. Provide interpretations of all of your answers. 

***

First, I'll define a function, `find_CI()` that finds the confidence interval, given a signifigance level, mean, sd, and n

```{r, echo=T, collapse=T}
# Define function to return CI based on normal/t distribution
find_CI <- function(SigLvl, x, n, s, population=T) {
  p <- SigLvl + (1 - SigLvl)/2
  
  # if using population sigma, use Z score 
  if(population) {
    error <- qnorm(p) * s/sqrt(n)
  } else({ # else use t-score
    error <- qt(p, df=n-1) * s/sqrt(n)
  })
  
  # Find CI's
  lower_ci <- x - error
  upper_ci <- x + error
  
  return( c(lower=lower_ci, upper=upper_ci) )
}
```

# Question 1

> 1. _A random sample of n=712 patient weighs is collected from a population of 20-29-year-old men and calculate a mean weight of $\bar{x}$=183 pounds. The population standard deviation, estimated from pilot data, is 40._

For this problem, we'll use the Z-score by setting `find_CI(..., , population=T)` because we know the population sd

## Find the 90% CI

> 1a) _Calculate a 90% confidence interval._



```{r, echo=T, collapse=T}
find_CI(SigLvl=0.90, x=183, n=712, s=40)
```

90% CI = [`r find_CI(SigLvl=0.90, x=183, n=712, s=40)[[1]]` to `r find_CI(SigLvl=0.90, x=183, n=712, s=40)[[2]]`]

## Find the 95% CI

> 1b) _Calculate a 95% confidence interval._

```{r, echo=T, collapse=T}
find_CI(SigLvl=0.95, x=183, n=712, s=40)
```

95% CI = [`r find_CI(SigLvl=0.95, x=183, n=712, s=40)[[1]]` to `r find_CI(SigLvl=0.95, x=183, n=712, s=40)[[2]]`]


## Comment on the difference between these two intervals.

> 1c) _Comment on the difference between these two intervals._

The 90% CI has a smaller range than the 95% CI, because it requires less confidence and has a smaller Z. Essentially, the higher confidence means a larger Z (to account for more area under the curve) which causes the range to be larger. See the table below as an example

```{r, echo=F}
tibble(CI_Lvl = c(.5, .80, .90, .95, .99, .999)) %>%
  group_by(CI_Lvl) %>%
  mutate(
    Z        = qnorm( CI_Lvl + (1 - CI_Lvl)/2 ),
    lower_ci = find_CI(CI_Lvl, x=183, n=712, s=40)[[1]],
    upper_ci = find_CI(CI_Lvl, x=183, n=712, s=40)[[2]],
    range    = upper_ci - lower_ci
  ) %>% pander()


# tibble(value = rnorm(172, mean=183, sd=40))

```



# Question 2

> _2. The data in Table 2.15 (p.37) present Serum-cholesterol levels (mg/dL) before and after adopting a vegetarian diet._

```{r}
Chol <- read_csv("Cholesterol.csv")

Chol %>% head() %>% pander(caption="First few obvs of Cholesterol.csv")
```


## Find the 95% CI

> 2a) _Compute a 95% confidence interval for the mean difference in Serum-cholesterol levels. _

**Null hypothesis (H0):** true difference in means equals 0      
**Alternative hypothesis (Ha):** true difference in means is not equal to 0

```{r, echo=T}
Chol <- read_csv("Cholesterol.csv")

t.test(Chol$Before, Chol$After, paired = T)
```

This gives us a **95% CI**  of [12.445 - 26.638]. Additional info from the test is in table below:

```{r}
pander(t.test(Chol$Before, Chol$After, paired = T))
```

The two other ways to calculate it below also give the same results:

```{r, echo=T, collapse=T}
# Same result with one-way t-test
t.test(Chol$Difference, var.equal = T)$conf.int

# Same result using my find_CI function
find_CI(SigLvl = .95, n = 24, 
        x = mean(Chol$Difference), 
        s = sd(Chol$Difference),
        population = F)
```







## When to use Z vs T

> 2b) _Explain when you should use the z or the t-distributions when constructing confidence intervals._

You use z-distributions when you know the population's standard deviation ($\sigma$). You can also use the z-distribution if your sample is large enough (e.g. n$\geq$ 30) as the Z and T socres begin to approximate each other.

```{r, eval=F, include=F}
tibble(
  t = qt(.975, df=1:100),
  z = qnorm(.975),
  x = 1:100
) %>%
  ggplot(aes(x=x, y=t)) +
  geom_line(color="red") +
  geom_hline(yintercept = qnorm(.975))

```



# Question 3

> 3. _Renal disease: The mean serum-creatinine level measured in 12 patients 24 hours after they received a newly proposed antibiotic was 1.2mg/dL._

## Test if difference in means 

> 3a) _If the mean and **standard deviation** of serum creatinine in the **general population** are 1.0 and 0.4 mg/dL, respectively, then, using a significance level of .05, test whether the mean serum-creatinine level in this group is different from that of the general population._

**H0:** true difference in means = 0     
**Ha:** true difference in means does not equal 0

```{r, echo=T, collapse=T}
# Define values
n <- 12
x_bar <- 1.2
mu    <- 1.0
s <- .4

# Find Z, since we know population sd
se <- s / sqrt(n)
z <- (x_bar - mu) / se

# p-value
2*pnorm(-abs(z))

# 95% CI
find_CI(.95, x=x_bar, n=n, s=s, population = T)
```

Given the results above, we **fail to reject the null hypothesis**


## p-value

> 3b) _What is the p-value for the test?_

As demonstrated above, the p-value for this test is `r round(2*pnorm(-abs(z)), 3)`

```{r, echo=T, collapse=T}
2*pnorm(-abs(z))
```


## Using t for the the same problem

> 3c) _Suppose the sample standard deviation of serum creatinine in Problem 3a is 0.6mg/dL. Assume that the standard deviation of serum creatinine is not known in the general population, and perform the hypothesis test in Problem 3a. Report a p-value._

For this problem we'll use the same hypotheses as before, but now with our `s = 0.6mg/dL`. This also means we'll need to do a t-test (since n=12).

```{r, echo=T, collapse=T}
s <- .6

# Find T
se <- s / sqrt(n)
t <- (x_bar - mu) / se

# p-value
2*pt(-abs(t),df=n-1)
```

Again, we fail to reject the null with a p = `r round(2*pt(-abs(t),df=n-1), 3)`


## Report the CI

> 3d) _Compute a two-sided 95% CI for the true mean serum-creatinine level in Problem 3c._

```{r, echo=T, collapse=T}
# 95% CI
find_CI(.95, x=x_bar, n=n, s=s, population = FALSE)
```

95% CI = `r find_CI(.95, x=x_bar, n=n, s=s, population = F)[[1]]` to `r find_CI(.95, x=x_bar, n=n, s=s, population = F)[[2]]`


# Question 4

> 4. _The drug diflunisal is used to treat mild to moderate pain due to osteoarthritis (OA) and rheumatoid arthritis (RA). The ocular effects of diflunisal had not been considered until a study was conducted on its effect on intraocular pressure in glaucoma patients who were already receiving maximum therapy for glaucoma. _

## Two-sided t-test

> 4a) _The change in intraocular pressure after administration of diflunisal among 30 patients whose standard therapy was topical drugs only has a mean -0.7 and sample standard deviation 2.1 mm Hg. Does diflunisal has **any effect** on the intraocular pressure? State the hypothesis, perform the test and interpret._

**H0:** true difference in means = 0     
**Ha:** true difference in means does not equal 0

```{r, echo=T, collapse=T}
n <- 30
x_bar <- -0.7
s <- 2.1

# t value
t <- x_bar / (s / sqrt(n))
t

# p-value
p <- 2*pt(-abs(t),df=n-1)
p
```

This tests yields t=`r round(t, 4)` on df=`r n-1`, with a resulting p = `r round(p,4)`. We therefore **fail to reject the null hypothesis**.

## 95% CI

> 4b) _Compute the 95% CI for the mean change in pressure in 8.20. Interpret._

```{r, echo=T, collapse=T}
find_CI(.95, x=-.7, n=30, s=2.1, population = FALSE)
```

95% CI = `r find_CI(.95, x=-.7, n=30, s=2.1, population = FALSE)[[1]]` - `r find_CI(.95, x=-.7, n=30, s=2.1, population = FALSE)[[1]]`.  
This means that across repeated samples of size `n`, we would expect that around 95% of those confidence intervals would contain the population mean.

## One-sided t-test           

> 4c) _Does diflunisal **reduce** intraocular pressure? State the hypothesis, perform the test and interpret. (Compare it to the results in 4a)_

**H0:** true difference in means $\leq$ 0     
**Ha:** true difference in means is greater than 0

```{r, echo=T, collapse=T}
# One sided p-value
p <- pt(t,df=n-1)
p
```

In this case, the p-value (p=`r round(p,4)`) is half the size of our p-value in problem 4a. Unlike in 4a), if we use these null & alternative hypotheses, we can **reject this null hypothesis** (true difference in means $\leq$ 0) at the signifigance level alpha=0.05.


