---
title: "NSFG data clean"
author: "Hunter Ratliff"
date: "11/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pander)
```

## preg data

Here are the changes I've made:


I added these variables:

- **BirthWt:** if they refused (98) or didn't know (99) the variable __BIRTHWGT_LB1__, recode as `NA`. Note that there are already NA's for this variable (e.g. current pregnancy, abortion)
- **GA:** same as above
- **BeginPNcare:** same as above
- **KnowPreg:** same as above

```{r read_preg, message=F, warning=F}
preg <- read_csv("~/SASUniversityEdition/myfolders/NSFG/csv/preg.csv") %>%
  mutate(
    BirthWt     = ifelse(BIRTHWGT_LB1>95, NA, BIRTHWGT_LB1),
    GA          = ifelse(WKSGEST>94, NA, WKSGEST),
    BeginPNcare = ifelse(BGNPRENA>94, NA, BGNPRENA),
    KnowPreg    = ifelse(KNEWPREG>94, NA, KNEWPREG)
  ) %>%
  mutate(
    PMARPREG = recode(str_to_lower(PMARPREG), 
                      "no"=F, 
                      "yes"=T),
    LBW1     = recode(LBW1, 
                      "NO, NOT LOW BIRTH WEIGHT"=F,
                      "YES, LOW BIRTH WEIGHT"=T)
  ) %>%
  mutate_if(is.character, factor) %>%
  
  filter(
    GA < 45
  )
  
preg %>%
  summary() %>% 
  pander(justify="left")
```


# Resp

```{r read_resp, message=F, warning=F}
resp <- read_csv("~/SASUniversityEdition/myfolders/NSFG/csv/resp.csv") %>%
  mutate(BMI = ifelse(BMI>94, NA, BMI)) %>%

  mutate(
    ASKSMOKE = recode(ASKSMOKE, "No"=F, "Yes"=T, "Don't know"=NA),
    ASKPREG  = recode(ASKPREG, "No"=F, "Yes"=T, "Don't know"=NA),
    ASKFOLIC = recode(ASKFOLIC, "No"=F, "Yes"=T, "Don't know"=NA),
    USUALCAR = recode(USUALCAR, "No"=F, "Yes"=T, "Don't know"=NA),
    
    GENHEALT = na_if(GENHEALT, "Don't know"),
    GENHEALT = na_if(GENHEALT, "Refused"),
    GENHEALT = na_if(GENHEALT, "Not ascertained"),
    GENHEALT = factor(GENHEALT, ordered = T,
                      levels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
    
    HISPRACE = recode(HISPRACE, "Non-Hispanic Black"="Black",
                      "Non-Hispanic White"="White",
                      "Non-Hispanic Other"="Other"),
    
    PREG_INT1 = na_if(PREG_INT1, "Could not be defined"),
    PREG_INT1 = str_replace(PREG_INT1, "months", "mo"),
    PREG_INT1 = recode(PREG_INT1, "60 mo or longer"="60+ mo"),
    PREG_INT1 = factor(PREG_INT1, ordered = T,
                       c("0-5 mo", "6-11 mo", "12-17 mo", "18-23 mo", "24-59 mo", "60+ mo"))
    
  ) %>%
  mutate_if(is.character, factor) 

resp %>%
  # count(HIEDUC)
  summary() %>% 
  pander(justify="left")

  # glimpse()
```


# Simple descriptive stats

```{r}
glimpse(preg)
```


# Relationship between tables

The `preg` table has `r nrow(preg)` rows across `r length(unique(preg$CASEID))` unique subjects. This means that many women had more than one pregnancy, and the table below shows the breakdown of number of women by number of pregnancies

```{r collapse=T, echo=F}
preg %>%
  count(CASEID, sort = T) %>%
  rename(NumPregs=n) %>%
  count(NumPregs) %>%
  pander()
```

There are also `r nrow(anti_join(resp, preg))` records contained in `resp` that don't have corresponding rows in `preg`



# Ignore

```{r include=F, eval=F}
# Just for fun
library(randomForest)
require(caTools)
df <- preg %>% 
  filter(OUTCOME!="CURRENT PREGNANCY") %>%
  filter(!is.na(OUTCOME), !is.na(GA)) %>%
  mutate_if(is.character, factor)

colSums(is.na(df))

# Partion data
sample <- sample.split(df$OUTCOME, SplitRatio = .75)
train <- subset(df, sample == TRUE)
test  <- subset(df, sample == FALSE)

# Random forest
rf <- randomForest(
  OUTCOME ~ AGECON + PREGORDR + WANTRESP + GA,
  # proximity = T,
  data=train
)
rf
varImpPlot(rf)
getTree(rf, k = 1, labelVar = T)
plot(rf)
varUsed(rf)
rf %>% margin() %>% plot()


# MDSplot(rf, train$OUTCOME)
rm(df, sample, train, test, rf)
```


