---
title: "Some models"
author: "Hunter Ratliff"
date: "11/17/2019"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    df_print: kable
    highlight: tango
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F, collapse = T)
library(pander)
library(scales)
library(ggthemes)
# library(randomForest)
# library(survey) # dsgn <- svydesign(ids=~CASEID, strata = ~strata, weights=~wgt, data=data)
# require(caTools)
# library(caret)
library(leaps)
library(MASS)
library(kableExtra)
library(tidyverse)
# library(Epi)        # Epi::clogistic for conditional logit
# library(survival)   # survival::clogit()
### survival allows for weights
```

# Set reference levels

The reference levels can be a little confusing for the regression, so I'll explain it here.

For predictors (i.e. _independent variables_), the reference is exactly what it sound like. So if I set the reference level for `race` to be white, the estimate for coefficients race`XYZ` is in reference to white subjects, as shown below:

```{r, eval=FALSE, echo=T}
BIRTH %>%
  mutate(
    race = relevel(race, "White")
  )

# Coefficients:
#                Estimate Std. Error z value Pr(>|z|)    
# ...              ...        ...       ...   ...
# raceHispanic  3.476e-02     ...       ...   ...
```

For our our _outcomes_ it's the **second** level of the factor that becomes our outcome (or "success"). So if the reference is set to having a normal birthweight, then the odds ratio is representing the odds ratio of **being born low birth weight**.


```{r readData}
# Read in data
BIRTH <- read_rds("data/BIRTH.RDS") %>%
  mutate(
    LBW = factor(ifelse(LBW, "Low", "Normal")),
    Wanted = factor(ifelse(Wanted, "Yes", "No")),
    
    ### --------------- Set default levels ---------------
    ## For OUTCOME variables - see help(binomial)
    # For using factors in binomial: 'success' is interpreted as the factor 
    # NOT having the first level (success=second level)
    LBW   = relevel(LBW,   "Normal"),
    PreMe = relevel(PreMe, "Term"),
    
    ## For PREDICTORS, the first level is the ref
    Wanted = relevel(Wanted, "Yes"),
    race   = relevel(race,   "White")
  )

PN    <- read_rds("data/PN.RDS") %>%
  mutate(
    LBW = factor(ifelse(LBW, "Low", "Normal")),
    Wanted = factor(ifelse(Wanted, "Yes", "No")),
    
    ### --------------- Set default levels ---------------
    ## For OUTCOME variables - see help(binomial)
    # For using factors in binomial: 'success' is interpreted as the factor 
    # NOT having the first level (success=second level)
    LBW   = relevel(LBW,   "Normal"),
    PreMe = relevel(PreMe, "Term"),
    
    ## For PREDICTORS, the first level is the ref
    Wanted = relevel(Wanted, "Yes"),
    race   = relevel(race,   "White")
  )
```


Below are the reference levels for the `BIRTH` dataset

```{r}
BIRTH %>% select_if(is.factor) %>% select(-eduCat) %>% str()
```

Below are the reference levels for the `PN` dataset

```{r}
PN %>% select_if(is.factor) %>% select(-eduCat) %>% str()
```


# Odds ratio based on 2x2 tables

Odds ratio of being LBW, given exposure to unwantedness

```{r OR_example, echo=T, collapse=T}
tab <- t(table(BIRTH$LBW, BIRTH$Wanted)) 
tab
tab %>% mosaic::orrr()  %>% .^-1
# Odds ratio of being LBW, given exposure to unwantedness
# Odds LBW in unwanted: 222/1834 = .12105
# Odds LBW in wanted: 236/2528 = .09335
(222/1834)/(236/2528) # 1.296638
glm(LBW ~ Wanted, data=BIRTH, family=binomial) %>% coef() %>% exp()
# glm(LBW ~ Wanted, data=BIRTH, family=binomial) %>% 
#   predict(newdata = tibble(PreMe=as.factor("Term")), type="response")
```

So we see that we should be interpreting the odds ratio of the regression as follows

> the odds of [**second level of outcome variable**], given exposure to [**second level of predictor**]

In this case, the odds ratio shows us the odds of *being __low__ birth weight*, with a reference level of *wanted pregnancy* (i.e. OR of being LBW given unwanted pregnancy)

# Low Birth Weight

First, a refresher on the number of subjects who were low birth weight

```{r}
table(BIRTH$LBW) %>% 
  pander(caption="Counts of subjects with normal and low birth weights")
```

This model predicts being born as **low birth weight** with the formula below

```{r}
mod.LBW <- glm(LBW ~ GA + BMI + age + income + race + YrEdu + Wanted, 
    data=BIRTH, family=binomial(link = "logit"))
pander(mod.LBW)

# Anova(mod.LBW, type = 3)  # type 3 anova
```

Note that the estimate above is the **Log-odds**. The table below transforms the log-odds into odds ratios

```{r}
# Tidy version of summary
broom::tidy(mod.LBW) %>%
  select(term, estimate, p.value) %>%
  mutate(Odds.Ratio = exp(estimate)) %>%
  select(-estimate) %>%
  cbind(exp(broom::confint_tidy(mod.LBW))) %>%
  mutate_if(is.numeric, round, 4) %>%
  filter(term!="(Intercept)") %>%
  mutate(
    p.value    = cell_spec(p.value, color = ifelse(p.value <0.05, "red", "black"))
    ) %>%
  knitr::kable(digits=4, escape=F) %>%
  kable_styling("striped", full_width = F)
# Also can do this --> exp(cbind(OR = coef(mod.LBW), confint(mod.LBW))) %>% round(3)    

```

## Graphics

```{r prediction1}
### Predictions of propabilities ###
df <- tidyr::expand(BIRTH, Wanted, race) %>%
  mutate(
    GA = mean(BIRTH$GA),
    BMI = mean(BIRTH$BMI),
    age = mean(BIRTH$age),
    # income = mean(BIRTH$income),
    YrEdu  = mean(BIRTH$YrEdu)
  ) %>%
  expand_grid(
    income=seq(0, 500, by=50)
  ) 

# predict(object = mod.LBW, newdata = df, type="response", se.fit = T)
df$fit <- predict(object = mod.LBW, newdata = df, type="response", se.fit = T)$fit
df$ul  <- df$fit + predict(object = mod.LBW, newdata = df, type="response", se.fit = T)$se.fit
df$ll  <- df$fit - predict(object = mod.LBW, newdata = df, type="response", se.fit = T)$se.fit

df %>% ggplot(aes(x=income, y=fit, group=race)) +
  geom_ribbon(aes(ymax=ul, ymin=ll), alpha=0.1) +
  geom_line(aes(color=race)) +
  facet_grid(.~Wanted) +
  scale_color_colorblind() +
  theme_bw() +
  labs(x="Income as % federal poverty line", y="Pr(Having LBW baby)",
       title="Predicted Pr(LBW), by income & race",
       subtitle = "Split by Wantedness (Yes/No)")
```

```{r prediction2}
### Predictions of propabilities ###
df <- tidyr::expand(BIRTH, race) %>%
  mutate(
    # GA = mean(BIRTH$GA),
    BMI = mean(BIRTH$BMI),
    age = mean(BIRTH$age),
    income = mean(BIRTH$income),
    Wanted = as.factor("Yes"),
    YrEdu  = mean(BIRTH$YrEdu)
  ) %>%
  expand_grid(
    GA=seq(20, 45, by=1)
  ) 

# predict(object = mod.LBW, newdata = df, type="response", se.fit = T)
df$fit <- predict(object = mod.LBW, newdata = df, type="response", se.fit = T)$fit
df$ul  <- df$fit + predict(object = mod.LBW, newdata = df, type="response", se.fit = T)$se.fit
df$ll  <- df$fit - predict(object = mod.LBW, newdata = df, type="response", se.fit = T)$se.fit

df %>% ggplot(aes(x=GA, y=fit, group=race)) +
  geom_ribbon(aes(ymax=ul, ymin=ll), alpha=0.1) +
  geom_line(aes(color=race)) +
  # facet_grid(.~Wanted) +
  scale_color_colorblind() +
  theme_bw() +
  labs(x="Gestational age (weeks)", y="Pr(Having LBW baby)",
       title="Predicted Pr(LBW), by gestational age & race")
```

## Validation

VIF 

```{r}
car::vif(mod.LBW)
```

Confusion matrix using test & training data

```{r}
train <- BIRTH %>% group_by(LBW) %>% sample_frac(.8, replace=T)
test <- anti_join(BIRTH, train)
# BIRTH %>% group_by(LBW) %>% sample_n(.8*nrow(BIRTH), replace=T) # gives equal numbers


mod <- glm(mod.LBW[["formula"]], data=train, family=binomial(link = "logit"))
predicted <- predict(mod, test, type="response")

library(InformationValue)
# find optimal cutoff
cutoff <- optimalCutoff(ifelse(test$LBW=="Low", 1, 0), predicted)

# find error rate
misClassError(ifelse(test$LBW=="Low", 1, 0), predicted,
              threshold = cutoff)

# plot ROC
plotROC(ifelse(test$LBW=="Low", 1, 0), predicted)

Concordance(ifelse(test$LBW=="Low", 1, 0), predicted)


# Confusion matrix
caret::confusionMatrix(positive="Low",
  as.factor(ifelse(predicted>cutoff, "Low", "Normal")),
  reference=test$LBW)

# clean up 
rm(train, test, mod, predicted, cutoff)
```



# Premature birth

Again, a refresher on the number of subjects who were born before 37 weeks

```{r}
table(BIRTH$PreMe) %>% 
  pander(caption="Counts of subjects born at term and prematurely")
```

This model predicts **being born prematurely** with the formula below

```{r}
mod.PreMe <- glm(PreMe ~  BMI + age + income + race + YrEdu + Wanted, 
    data=BIRTH, family=binomial(link = "logit"))
pander(mod.PreMe)
```

Again, table with the transformed odds ratios

```{r}
# Tidy version of summary
broom::tidy(mod.PreMe) %>%
  select(term, estimate, p.value) %>%
  mutate(Odds.Ratio = exp(estimate)) %>%
  select(-estimate) %>%
  cbind(exp(broom::confint_tidy(mod.PreMe))) %>%
  mutate_if(is.numeric, round, 4) %>%
  filter(term!="(Intercept)") %>%
  mutate(
    p.value    = cell_spec(p.value, color = ifelse(p.value <0.05, "red", "black"))
    ) %>%
  knitr::kable(digits=4, escape=F) %>%
  kable_styling("striped", full_width = F)
```




# Knew pregnant

Counts of who knew they were pregnant

```{r}
table(PN$KnowPreg) %>% 
  pander(caption="Counts of if mother knew pregnant or not by XXX weeks")
```

This model predicts ...

```{r}
mod.KnowPreg <- glm(KnowPreg~PregNum + age + income + race + YrEdu + PregNum + Wanted, 
    data=PN, family=binomial)
pander(mod.KnowPreg)
```

Odds ratios

```{r}
# Tidy version of summary
broom::tidy(mod.KnowPreg) %>%
  select(term, estimate, p.value) %>%
  mutate(Odds.Ratio = exp(estimate)) %>%
  select(-estimate) %>%
  cbind(exp(broom::confint_tidy(mod.KnowPreg))) %>%
  mutate_if(is.numeric, round, 4) %>%
  filter(term!="(Intercept)") %>%
  mutate(
    p.value    = cell_spec(p.value, color = ifelse(p.value <0.05, "red", "black"))
    ) %>%
  knitr::kable(digits=4, escape=F) %>%
  kable_styling("striped", full_width = F)
```


# Got prenatal care

Counts of who got prenatal care in the first trimester

```{r}
table(PN$gotPNcare) %>% 
  pander(caption="Counts of if mother got prenatal care in the first trimester")
```

```{r}
mod.gotPNcare <- glm(gotPNcare~ KnowPreg + age + income + race + YrEdu + PregNum + Wanted, 
    data=PN, family=binomial)
pander(mod.gotPNcare)
```

Odds ratios

```{r}
# Tidy version of summary
broom::tidy(mod.gotPNcare) %>%
  select(term, estimate, p.value) %>%
  mutate(Odds.Ratio = exp(estimate)) %>%
  select(-estimate) %>%
  cbind(exp(broom::confint_tidy(mod.gotPNcare))) %>%
  mutate_if(is.numeric, round, 4) %>%
  filter(term!="(Intercept)") %>%
  mutate(
    p.value    = cell_spec(p.value, color = ifelse(p.value <0.05, "red", "black"))
    ) %>%
  knitr::kable(digits=4, escape=F) %>%
  kable_styling("striped", full_width = F)
```




